---
title: "Project Part 3"
output: 
  rmdformats::robobook:
    css: "homework.css"
    highlight: pygments
link-citations: yes
---

**Due dates:** Friday, December 5 at 11:59pm on Canvas.

**Submission:** A template file is provided, with a rough outline: [project_1_3_template.qmd](https://sta214-f25.github.io/projects/project_1_3_template.qmd)

Once you have completed the report, submit *both* your 
rendered HTML file and the Quarto file to Canvas.

# Overview

This semester, you will complete a project in which you use logistic regression to investigate a research question of your choosing. You will then write a report that communicates the results of your analysis and draws a conclusion about your research question. 

In the final part of the project, you will combine your work from the previous parts to create a full report.


# Your task in Part III

You will submit two documents on Canvas for this project.

* Your written report (.html file): 
  * This is the knitted write up that will explain your work, and address your research question
  * There should be **no code** showing. This
includes warnings and other stray code output – hide it all.
  * You will be graded on writing as well as your statistical analysis.
* Your Quarto (.qmd) file, with all the code you used for your analysis
  * The goal is that a person who reads your report, and wants to
replicate your results, could look at your .qmd file and completely reproduce the results and figures in the rendered report.

Each section should be written in paragraph form, and make sure to always discuss any figures in your report.

There is no minimum or maximum length required. However, your analysis should be thorough, with your report addressing the research question, and containing all the elements described below.


# Content

Your report should contain the following labeled sections.

## Abstract

This is the first section in your report, but it is actually the last thing you will write. Called an abstract in academia, and an executive summary in industry, this one paragraph summary of your entire paper is the first thing people will read. For example:

*Estimating the number of pandas inhabiting a national panda preserve is critical to understanding the health of the panda species as a whole. In this report, we discuss the process of estimating the number of pandas in the Wolong National Nature Reserve. Data were collected from volunteers walking trails in the park, and we applied capture-recapture estimation process to then estimate the total number of pandas in the park. We detail the steps of the process, and compare our results to two other possible methods of estimating pandas. Based on our estimation, there are between 140 and 155 pandas in the surveyed area. We discuss our findings, and limitations of the study, in the following paper.*
  
## Section 1: Introduction

Write 2-3 paragraphs in which you:

* Motivate the research questions
* Describe where the data came from
* Provide background on the dataset
* Finish by summarizing what you will do in the report

## Section 2: Data

In this section, you will summarize and explore your data. First, write 1-2 paragraphs which:

* Describe details of the data. What does a row in the data represent? How many rows and columns? What information do the variables record? Are there any missing data?
* Describe any data manipulation. Did you remove any missing values? Did you focus only on a subset of the data? Did you create any new variables?

Next, perform exploratory data analysis. Your goal is to create any visualizations necessary to exploring the data before fitting a model (e.g. tables, histograms, empirical log-odds plots to choose transformations, etc.). Make sure that:

* You create a nicely formatted table to show the counts of your response variable
* You have at least three graphs shown
* Your figures have axis labels, captions, and are numbered Figure 2.1, Figure 2.2, etc., where the 2. represents the section number, and the number after the decimal represents how many plots so far in this section
* Any plot that is shown should be discussed in writing in your report, and referenced by number (e.g., "As shown in Figure 2.1...").  What information does this graph give us, and why is that important for us to know when before we start building our model in the next section?

## Section 3: Modeling

In this section, you will build a model to answer the research question. Here you will use your *final* regression model (chosen after making any adjustments to satisfy regression assumptions, such as variable transformations). Other preliminary or intermediate models which you considered can be included and discussed in the Appendix (below).

### Section 3.1: Model

First, we need to describe the regression model and give the fitted model. Using your *final* regression model, write a paragraph in which you:

* Explain your choice of model and variables
* Present a nicely formatted table of coefficients for your fitted model (see the section on Tables below)
* Interpret any relevant fitted coefficients in context of the research question. It may help to consider discussing coefficients in terms of a one standard deviation increase in the explanatory variable, rather than a one-unit increase.

### Section 3.2: Diagnostics

Next, we want to verify that the regression assumptions are reasonable for our model. Write a couple paragraphs in which you:

* Create and discuss quantile residual plots to assess the shape assumption
* Check for collinearity and influential points

### Section 3.3: Inference

Finally, we want to use our fitted regression model to address the research question. Write a couple paragraphs in which you perform a hypothesis test that allows you to address your research question. When testing hypotheses, make sure to:

* State the hypotheses in terms of one or more model parameters
* Calculate a test statistic and p-value
* Use the p-value to make a conclusion about the research question

## Section 4: Discussion

In this section, you are providing a conclusion. Write 1–2 paragraphs which summarize what you learned from your analysis, and how it addresses the original research questions. Also discuss any limitations to your analysis, and what you other relationships you might explore in future.


## Sources

Provide information on any sources you have used outside the course materials.

* Cite the original source of your data
* Cite any articles you have referenced
* Provide links to any websites you have used for code help
* If generative AI (Chat GPT, Gemini, etc.) was used to help with any part of the assignment, summarize
    * The AI tool used
    * The prompts you used
    * The help provided by the tool
    
## Appendix

If you considered any other models (other than your final model), or created any other exploratory or diagnostic plots that were not included in your main report, you can include these here.


# Appearance and style

### Writing

The report should be written like an article or research paper: in full sentences and paragraphs, with headings for each section. You should not write your report with question numbers or as a list of bullet points. Scientific articles are generally written in third person, though "we" can also be acceptable ("we can see from Figure 2.1 ...") in some disciplines.

### Code

As with previous assignments, you will write the document in Quarto. However, in full reports, the only output that should be visible from code chunks are figures and tables. If a code chunk does not produce a figure or table, you can hide it from the knitted document with `include=F`:

````
```{r, include=F}`r ''`

```
````

If a code chunk produces a figure or table, only the figure or table should be visible in the knitted document. You can hide the chunk but display the output with `echo=F, message=F, warning=F`:

````
```{r, echo=F, message=F, warning=F}`r ''`

```
````

### Figures

Figures should have labeled axes, and should be clear and easy to read. Figures should also be captioned and numbered; to caption a figure, use `fig.cap = "..."` in the chunk options. For example (scroll to the right on the code to see it all),

````
```{r, echo=F, message=F, warning=F, fig.cap="Figure 2.1: Bill depth vs. bill length for penguins near Palmer Station, Antarctica."}`r ''`
penguins |>
  ggplot(aes(x = bill_length_mm, 
             y = bill_depth_mm)) +
  geom_point() +
  labs(x = "Bill length (mm)",
       y = "Bill depth (mm)") +
  theme_bw()
```
````

is displayed as

```{r include=F}
library(tidyverse)
library(palmerpenguins)
```

```{r, echo=F, message=F, warning=F, fig.cap="Figure 1: Bill depth vs. bill length for penguins near Palmer Station, Antarctica."}
penguins |>
  ggplot(aes(x = bill_length_mm, 
             y = bill_depth_mm)) +
  geom_point() +
  labs(x = "Bill length (mm)",
       y = "Bill depth (mm)") +
  theme_bw()
```

Captions should provide enough information to understand what is being plotted, but interpretation can be left to the main text. Refer to figures by their number in the text. Make sure that any figures you include are discussed in the text.

### Tables

Tables should be nicely formatted, and have a number and caption. This can be done with the `kable` function in the `knitr` package.

For example,

````
```{r, echo=F, message=F}`r''`
table(penguins$island, penguins$species) |>
  knitr::kable(caption = "Table 3.2: Penguins by island and species")
```
````

is displayed as

```{r, echo=F, message=F}
table(penguins$island, penguins$species) |>
  knitr::kable(caption = "Table 3.2: Penguins by island and species")
```

### Writing math in Quarto or R Markdown

If you want to write mathematical notation, we need to tell Markdown, “Hey, we’re going to make a math symbol!” To do that, you use dollar signs. For instance, to make $\widehat{\beta}_1$, you simply put `$\widehathat{\beta}_1$` into the white space (not a chunk) in youQuarto.

Here are some examples of writing math, which you can adapt:

| Math | Code |
| --- | --- |
| $Y_i \sim Bernoulli(\pi_i)$ | `$Y_i \sim Bernoulli(\pi_i)$` |
| $\log \left( \dfrac{\pi_i}{1 - \pi_i} \right)$ | `$\log \left( \dfrac{\pi_i}{1 - \pi_i} \right)$` |
| $\widehat{\pi}_i$ | `$\widehat{\pi}_i$` |

