---
title: Inference
output:
  xaringan::moon_reader:
    css: "lab-slides.css"
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
editor_options: 
  chunk_output_type: console
---

### Last time

```{r message=F, warning=F, include=F}
library(tidyverse)
```


.large[
$$
\begin{aligned}
\text{Admit}_i &\sim Bernoulli(\pi_i) \\
\log \left( \dfrac{\pi_i}{1 - \pi_i} \right) &= \beta_0 + \beta_1 \text{GPA}_i + \beta_2 \text{GRE}_i + \\
& \hspace{0.5cm} \beta_3 \text{Rank2}_i + \beta_4 \text{Rank3}_i + \beta_5 \text{Rank4}_i
\end{aligned}
$$

**Research question:** Is there a relationship between GRE and grad school acceptance, after accounting for GPA and school rank?

$H_0: \beta_2 = 0 \hspace{1cm} H_A: \beta_2 \neq 0$
]

---

### A new research question

.large[
$$
\begin{aligned}
\text{Admit}_i &\sim Bernoulli(\pi_i) \\
\log \left( \dfrac{\pi_i}{1 - \pi_i} \right) &= \beta_0 + \beta_1 \text{GPA}_i + \beta_2 \text{GRE}_i + \\
& \hspace{0.5cm} \beta_3 \text{Rank2}_i + \beta_4 \text{Rank3}_i + \beta_5 \text{Rank4}_i
\end{aligned}
$$

**Research question:** Is there a relationship between school rank and grad school acceptance, after accounting for GPA and GRE score?

.question[
What are my hypotheses for this new research question?
]
]

---

### Testing multiple parameters

.large[
$$
\begin{aligned}
\log \left( \dfrac{\pi_i}{1 - \pi_i} \right) &= \beta_0 + \beta_1 \text{GPA}_i + \beta_2 \text{GRE}_i + \\
& \hspace{0.5cm} \beta_3 \text{Rank2}_i + \beta_4 \text{Rank3}_i + \beta_5 \text{Rank4}_i
\end{aligned}
$$

$H_0: \beta_3 = \beta_4 = \beta_4 = 0$

$H_A: \text{ at least one of } \beta_3, \beta_4, \beta_5 \neq 0$

.question[
How do we test multiple parameters in a *linear* regression model?
]
]

---

### Nested models

.large[
$H_0: \beta_3 = \beta_4 = \beta_4 = 0$

$H_A: \text{ at least one of } \beta_3, \beta_4, \beta_5 \neq 0$

**Full model** ($H_A$):

$$
\begin{aligned}
\log \left( \dfrac{\pi_i}{1 - \pi_i} \right) &= \beta_0 + \beta_1 \text{GPA}_i + \beta_2 \text{GRE}_i + \\
& \hspace{0.5cm} \beta_3 \text{Rank2}_i + \beta_4 \text{Rank3}_i + \beta_5 \text{Rank4}_i
\end{aligned}
$$

**Reduced model** ($H_0$):
]

---

### Comparing nested models

.large[
**Full model** ($H_A$):

$$
\begin{aligned}
\log \left( \dfrac{\pi_i}{1 - \pi_i} \right) &= \beta_0 + \beta_1 \text{GPA}_i + \beta_2 \text{GRE}_i + \\
& \hspace{0.5cm} \beta_3 \text{Rank2}_i + \beta_4 \text{Rank3}_i + \beta_5 \text{Rank4}_i
\end{aligned}
$$

**Reduced model** ($H_0$):

$$
\begin{aligned}
\log \left( \dfrac{\pi_i}{1 - \pi_i} \right) &= \beta_0 + \beta_1 \text{GPA}_i + \beta_2 \text{GRE}_i
\end{aligned}
$$
.question[
How should we compare these two nested models?
]
]

---

### Comparing nested models

.large[
**Deviance:** for a logistic regression model, deviance is $-2 \log L$

**Idea:**

* Maximizing likelihood (how we fit the model) is equivalent to minimizing deviance
* Smaller deviance implies better fit
* Is there a big difference in the deviance (model fit) between these models??

**Change in deviance:**

]

---

### Comparing deviance

```{r, include=F}
grad_app <- read.csv("https://stats.idre.ucla.edu/stat/data/binary.csv")
```


.large[
```{r}
m_full <- glm(admit ~ gpa + gre + as.factor(rank),
              family = binomial, data = grad_app)

m_reduced <- glm(admit ~ gpa + gre,
                 family = binomial, data = grad_app)

m_reduced$deviance - m_full$deviance
```

.question[
How do I know whether this is a big change in deviance? That is: **is a change in deviance of 21.83 unusual under the null hypothesis?**
]
]

---

### Parametric bootstrap

---

### Class activity

.large[
* Work on the class activity (linked on Canvas and course website)
* Submit rendered HTML on Canvas at the end of class
]

---

### Activity

```{r param_boot, cache=T, echo=F, fig.width=8, fig.height=5}
m_reduced <- glm(admit ~ gpa + gre,
                 family = binomial, data = grad_app)

nsim <- 5000
parametric_boot_stats <- rep(NA, nsim)

for(i in 1:nsim){
  # create a new parametric bootstrap sample
  # (code from question 1)
  admit_star <- rbinom(n = length(m_reduced$y), 
                     size = 1,
                     p = m_reduced$fitted.values)

  sim_data <- grad_app |>
    mutate(admit_star = admit_star)
  
  # re-fit the full and reduced models on the 
  # simulated data
  # (code from questions 2 - 3)
  m_full_sim <- glm(admit_star ~ gpa + gre + as.factor(rank),
                  family = binomial, data = sim_data)
  
  m_reduced_sim <- glm(admit_star ~ gpa + gre,
                  family = binomial, data = sim_data)
  
  # recompute the change in deviance, and save
  parametric_boot_stats[i] <- m_reduced_sim$deviance -
    m_full_sim$deviance
}
```

.large[
.question[
What does this distribution look like?
]
]

---

### Activity

```{r echo=F, fig.width=8, fig.height=5}
hist(parametric_boot_stats)
```

.large[
.question[
How unusual is the observed test statistic of 21.83, compared to this null distribution?
]
]

---

### p-value

```{r echo=F, fig.width=8, fig.height=5}
hist(parametric_boot_stats)
```

.large[
```{r}
mean(parametric_boot_stats > 21.83)
```
]





