---
author: "Ciaran Evans"
title: Logistic regression interpretation
output:
  xaringan::moon_reader:
    css: "lab-slides.css"
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
editor_options: 
  chunk_output_type: console
---

```{r include=F}
dengue <- read.csv("https://sta279-s22.github.io/labs/dengue.csv")
```


### Logistic regression model

.large[

$Y_i =$ dengue status (0 = negative, 1 = positive) 

$Age_i =$ age (in years)

**Random component:** $\hspace{1cm} Y_i \sim Bernoulli(\pi_i)$

**Systematic component:** $\hspace{1cm} \log \left( \dfrac{\pi_i}{1 - \pi_i} \right) = \beta_0 + \beta_1 \ Age_i$

]

---

### Fitting the logistic regression model

.large[
$$\hspace{1cm} Y_i \sim Bernoulli(\pi_i)$$
$$\hspace{1cm} \log \left( \dfrac{\pi_i}{1 - \pi_i} \right) = \beta_0 + \beta_1 \ Age_i$$

```{r, include=F}
library(knitr)
hook_output <- knit_hooks$get("output")
knit_hooks$set(output = function(x, options) {
   lines <- options$output.lines
   if (is.null(lines)) {
     return(hook_output(x, options))  # pass to default hook
   }
   x <- unlist(strsplit(x, "\n"))
   more <- "..."
   if (length(lines)==1) {        # first n lines
     if (length(x) > lines) {
       # truncate the output, but add ....
       x <- c(head(x, lines), more)
     }
   } else {
     x <- c(more, x[lines], more)
   }
   # paste these lines together
   x <- paste(c(x, ""), collapse = "\n")
   hook_output(x, options)
 })
```

```{r, eval=F}
m1 <- glm(Dengue ~ Age, data = dengue, 
          family = binomial)
summary(m1)
```
]

---

### Fitting the logistic regression model

.large[
$$\hspace{1cm} Y_i \sim Bernoulli(\pi_i)$$
$$\hspace{1cm} \log \left( \dfrac{\pi_i}{1 - \pi_i} \right) = \beta_0 + \beta_1 \ Age_i$$

```{r, output.lines = 5:8}
m1 <- glm(Dengue ~ Age, data = dengue, 
          family = binomial)
summary(m1)
```
]

---

### Last time: Activity

.large[
$$\hspace{1cm} \log \left( \dfrac{\widehat{\pi}_i}{1 - \widehat{\pi}_i} \right) = -2.45 + 0.22 \ Age_i$$

.question[
Does the probability of dengue increase or decrease with age?
]
]

---

### Activity

.large[
$$\hspace{1cm} \log \left( \dfrac{\widehat{\pi}_i}{1 - \widehat{\pi}_i} \right) = -2.45 + 0.22 \ Age_i$$

.question[
What are the predicted odds of dengue for a 5 year old patient?
]
]

---

### Activity

.large[
$$\hspace{1cm} \log \left( \dfrac{\widehat{\pi}_i}{1 - \widehat{\pi}_i} \right) = -2.45 + 0.22 \ Age_i$$

.question[
For a 5 year old patient: is the predicted *probability* of dengue $> 0.5$, $< 0.5$, or $= 0.5$?
]
]

---

### Calculating probability

.large[
$$\hspace{1cm} \log \left( \dfrac{\widehat{\pi}_i}{1 - \widehat{\pi}_i} \right) = -2.45 + 0.22 \ Age_i$$

Predicted *probability* of dengue for a 5 year old patient?
]

---

### Activity

.large[
Work on the activity (handout) with a neighbor, then we will discuss as a class.
]

---


### Activity

.large[
$$\hspace{1cm} \log \left( \dfrac{\widehat{\pi}_i}{1 - \widehat{\pi}_i} \right) = -2.45 + 0.22 \ Age_i$$

.question[
What is the predicted probability of dengue for a 10 year old patient?
]
]

---

### Activity 

.large[
$$\hspace{1cm} \log \left( \dfrac{\widehat{\pi}_i}{1 - \widehat{\pi}_i} \right) = -2.45 + 0.22 \ Age_i$$

.question[
Suppose we want to identify patients for whom the predicted probability of dengue is at least 0.5. What age range should we focus on?
]
]

---

### Activity 

.large[
$$\hspace{1cm} \log \left( \dfrac{\widehat{\pi}_i}{1 - \widehat{\pi}_i} \right) = -2.45 + 0.22 \ Age_i$$

.question[
Compare the odds of dengue for a 12 year old patient to the odds of dengue for an 11 year old patient. What do you notice?
]
]

---

### Interpretation

.large[
$$\hspace{1cm} \log \left( \dfrac{\widehat{\pi}_i}{1 - \widehat{\pi}_i} \right) = -2.45 + 0.22 \ Age_i$$

**Slope:**

* A one-year increase in age is associated with an increase of 25% in the *odds* of dengue.

**Intercept:**
]

---

### Interpretation

.large[
$$\hspace{1cm} \log \left( \dfrac{\widehat{\pi}_i}{1 - \widehat{\pi}_i} \right) = -2.45 + 0.22 \ Age_i$$

**Slope:**

* A one-year increase in age is associated with an increase of 25% in the *odds* of dengue.

**Intercept:**

* when age is 0, the estimated odds of dengue is $e^{-2.45} = 0.086$
]

---

### Adding more variables

.large[
Now let's add WBC as a variable to the model:

```{r, eval=F}
m2 <- glm(Dengue ~ Age + WBC, data = dengue,
          family = binomial)
summary(m2)
```

$$\hspace{1cm} \log \left( \dfrac{\widehat{\pi}_i}{1 - \widehat{\pi}_i} \right) = 0.34 + 0.15 \ Age_i - 0.31 WBC_i$$

.question[
How should I interpret each coefficient in the fitted model?
]
]

---

### Adding more variables

.large[
What about a *categorical* predictor?

```{r, eval=F}
m3 <- glm(Dengue ~ Age + Abdominal, data = dengue,
          family = binomial)
summary(m3)
```

$$\hspace{1cm} \log \left( \dfrac{\widehat{\pi}_i}{1 - \widehat{\pi}_i} \right) = -0.25 + 0.22 \ Age_i + 0.21 Abdominal_i$$

.question[
How should I interpret the coefficient on abdominal pain?
]
]

---

### Interpretation

.large[
$$\hspace{1cm} \log \left( \dfrac{\widehat{\pi}_i}{1 - \widehat{\pi}_i} \right) = -2.45 + 0.22 \ Age_i$$

**Slope:**

* A one-year increase in age is associated with an increase of 25% in the *odds* of dengue.

.question[
Can we interpet the slope in terms of the the *probability* (rather than the odds or log odds)?
]
]

---

### For next week

.large[
**Reading:** in the textbook, read

* sections 6.7.1 - 6.7.3
* sections 2.2 - 2.4

]


