---
title: Inference
output:
  xaringan::moon_reader:
    css: "lab-slides.css"
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
editor_options: 
  chunk_output_type: console
---

### Motivation

```{r message=F, warning=F, echo=F, fig.width=6, fig.height=4}
library(tidyverse)

dengue <- read_csv("https://sta214-f26.github.io/homework/dengue.csv")

dengue |>
  ggplot(aes(x = as.factor(Dengue), y = WBC)) +
  geom_boxplot() +
  theme_bw() +
  labs(x = "Dengue")
```

.large[
It looks like there is a difference in WBC between positive and negative dengue cases.

.question[
How would we assess whether this is real, or due to chance?
]
]

---

### Motivation

.large[
Let $\mu_0$ be population mean WBC for non-dengue cases, and $\mu_1$ be population mean WBC for dengue cases.

**Want:** $\mu_1 - \mu_0$

**Estimate:** $$\overline{WBC}_1 - \overline{WBC}_0 = -4.36$$

.question[
How do we account for variability between samples??
]
]

---

### Confidence interval

.large[
A **confidence interval** gives a "reasonable range" of values for an unknown parameter.

**Want:** $\mu_1 - \mu_0$

**Estimate:** $$\overline{WBC}_1 - \overline{WBC}_0 = -4.36$$

**Confidence interval:**
]


---

### Bootstrap resampling

---

### Bootstrap resampling

.large[
A small subset of the original data:

```{r, echo=F}
set.seed(8234)
dengue_sub <- dengue |>
  select(WBC, PLT, Age, Dengue) |>
  sample_n(5, replace=F)

as.data.frame(dengue_sub |> sample_n(5, replace=F))
```

A bootstramp sample of these rows:

```{r, echo=F}
dengue_sub |>
  sample_n(5, replace=T) |>
  data.frame()
```



]

---

### Bootstrap resampling

```{r message=F, warning=F, echo=F, fig.width=10, fig.height=4}
library(patchwork)

set.seed(4)

p1 <- dengue |>
  ggplot(aes(x = as.factor(Dengue), y = WBC)) +
  geom_boxplot() +
  theme_bw() +
  labs(x = "Dengue")

dengue_boot <- dengue |>
  sample_n(size = n(), replace=T) # -4.47

p2 <- dengue_boot |>
  ggplot(aes(x = as.factor(Dengue), y = WBC)) +
  geom_boxplot() +
  theme_bw() +
  labs(x = "Dengue")

p1 + p2
```

---

### Activity

.large[
* Work on the class activity (linked on Canvas and course website)
* Submit rendered HTML on Canvas at the end of class
]

---

### Activity

.center[
```{r, echo=F, fig.width = 6, fig.height=4}
# desired number of bootstrap samples
n_boot = 1000

# create a vector in which to store the bootstrap samples
boot_ests <- rep(NA, n_boot)

# do the bootstrapping
for(i in 1:n_boot){
  # fill in the code to create a bootstrap sample 
  # (question 1)
  dengue_boot <- dengue |>
    sample_n(size = n(), replace=T)
  
  # fill in the code to re-calculate the difference in means
  # (question 3). We will store the difference in the
  # boot_ests vector
  boot_ests[i] <- dengue_boot |>
    summarize(mean_diff = mean(WBC[Dengue == 1]) - 
                          mean(WBC[Dengue == 0])) |>
    as.numeric()
}

# make a histogram of the results
hist(boot_ests)
```

.large[
.question[
Based on the bootstrap samples, what would be a reasonable range of values for the difference in means?
]
]
]

---

### Activity

.center[
```{r, echo=F, fig.width = 6, fig.height=4}
# make a histogram of the results
hist(boot_ests)
```

.large[
.question[
What does the distribution look like?
]
]
]

---

### Activity

.large[
Bootstrap percentile interval:

```{r}
quantile(boot_ests, c(0.025, 0.975))
```

Normal interval with bootstrap SE:

```{r}
-4.36 - 1.96*sd(boot_ests)
-4.36 + 1.96*sd(boot_ests)
```

]

---

### Activity

.large[
Comparison with the usual formula:

$$\widehat{SE}(\overline{X}_1 - \overline{X}_0) = \sqrt{\frac{s_1^2}{n_1} + \frac{s_0^2}{n_0}}$$
]

```{r}
sqrt(var(dengue$WBC[dengue$Dengue == 1])/sum(dengue$Dengue == 1) +
       var(dengue$WBC[dengue$Dengue == 0])/sum(dengue$Dengue == 0))
```

.large[
Bootstrap estimate of standard error:
]

```{r}
sd(boot_ests)
```


]

