---
title: Inference
output:
  xaringan::moon_reader:
    css: "lab-slides.css"
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
grad_app <- read.csv("https://stats.idre.ucla.edu/stat/data/binary.csv")

library(knitr)
hook_output <- knit_hooks$get("output")
knit_hooks$set(output = function(x, options) {
   lines <- options$output.lines
   if (is.null(lines)) {
     return(hook_output(x, options))  # pass to default hook
   }
   x <- unlist(strsplit(x, "\n"))
   more <- "..."
   if (length(lines)==1) {        # first n lines
     if (length(x) > lines) {
       # truncate the output, but add ....
       x <- c(head(x, lines), more)
     }
   } else {
     x <- c(more, x[lines], more)
   }
   # paste these lines together
   x <- paste(c(x, ""), collapse = "\n")
   hook_output(x, options)
 })
```

### Grad admissions data

.large[
Interested in predicting grad admission from GPA, GRE score, and school rank:

$$
\begin{aligned}
\text{Admit}_i &\sim Bernoulli(\pi_i) \\
\log \left( \dfrac{\pi_i}{1 - \pi_i} \right) &= \beta_0 + \beta_1 \text{GPA}_i + \beta_2 \text{GRE}_i + \\
& \hspace{0.5cm} \beta_3 \text{Rank2}_i + \beta_4 \text{Rank3}_i + \beta_5 \text{Rank4}_i
\end{aligned}
$$

.question[
Suppose we want to estimate the effect of GRE, after accounting for GPA and school rank. How would I compute a confidence interval for $\beta_2$?
]
]

---

### Previously: confidence intervals

.large[
Normal confidence interval for $\beta_2$:

$$\widehat{\beta}_2 \pm z^* SE_{\widehat{\beta}_2}$$

Options for estimating $SE_{\widehat{\beta}_2}$:

* Bootstrap estimate of standard error
* Estimate of standard error from R output
]

---

### Confidence interval

```{r, include=F}
m1 <- glm(admit ~ gpa + gre + as.factor(rank),
          family = binomial, data = grad_app)
summary(m1)
```

```
                     Estimate  Std. Error   z value     Pr(>|z|)
(Intercept)      -3.989979073 1.139950936 -3.500132 0.0004650273
gpa               0.804037549 0.331819298  2.423119 0.0153878974
gre               0.002264426 0.001093998  2.069864 0.0384651284
as.factor(rank)2 -0.675442928 0.316489661 -2.134171 0.0328288188
as.factor(rank)3 -1.340203916 0.345306418 -3.881202 0.0001039415
as.factor(rank)4 -1.551463677 0.417831633 -3.713131 0.0002047107
```

<br/>

<br/>

<br/>

.large[
.question[
How do I interpret this confidence interval?
]
]

---

### Another question

.large[
$$
\begin{aligned}
\text{Admit}_i &\sim Bernoulli(\pi_i) \\
\log \left( \dfrac{\pi_i}{1 - \pi_i} \right) &= \beta_0 + \beta_1 \text{GPA}_i + \beta_2 \text{GRE}_i + \\
& \hspace{0.5cm} \beta_3 \text{Rank2}_i + \beta_4 \text{Rank3}_i + \beta_5 \text{Rank4}_i
\end{aligned}
$$

**Research question:** Is there a relationship between GRE and grad school acceptance, after accounting for GPA and school rank?

.question[
How could we address this research question?
]
]

---

### Hypothesis testing

.large[
.question[
What steps are involved in hypothesis testing?
]
]

---

### Specifying the hypotheses

.large[
$$
\begin{aligned}
\log \left( \dfrac{\pi_i}{1 - \pi_i} \right) &= \beta_0 + \beta_1 \text{GPA}_i + \beta_2 \text{GRE}_i + \\
& \hspace{0.5cm} \beta_3 \text{Rank2}_i + \beta_4 \text{Rank3}_i + \beta_5 \text{Rank4}_i
\end{aligned}
$$

**Research question:** Is there a relationship between GRE and grad school acceptance, after accounting for GPA and school rank?

$H_0$:

<br/>

$H_A$:
]

---

### Confidence intervals and hypothesis testing

.large[
**Research question:** Is there a relationship between GRE and grad school acceptance, after accounting for GPA and school rank?

$H_0: \beta_2 = 0 \hspace{1cm} H_A: \beta_2 \neq 0$

Recall that we would **reject** $H_0$ at significance level $\alpha$ if and only if a $(1 - \alpha)100\%$ confidence interval for $\beta_2$ **does not** contain 0. (Equivalently, a $(1 - \alpha)100%$ confidence interval for $e^{\beta_2}$ **does not** contain 1).
]

---

### Computing a p-value

.large[
To compute a p-value, we need:

* a test statistic
* the distribution of the test statistic **if** $H_0$ **were correct**

.question[
What test statistic did we use for testing single coefficients in *linear* models?
]
]

---

### The null distribution

.large[
To compute a p-value, we need the **null distribution** (distribution of the test statistic under $H_0$).

.question[
Does *bootstrap resampling* (what we used to create a confidence interval) give us the **null** distribution?
]
]

---

### Parametric bootstrapping

---

### Class activity

.large[
* Work on the class activity (linked on Canvas and course website)
* Submit rendered HTML on Canvas at the end of class
]

---

### Activity

```{r param_boot, cache=T, echo=F, fig.width=8, fig.height=5}
m_full <- glm(admit ~ gpa + gre + as.factor(rank),
              family = binomial, data = grad_app)

m_reduced <- glm(admit ~ gpa + as.factor(rank),
                 family = binomial, data = grad_app)

nsim <- 5000
parametric_boot_stats <- rep(NA, nsim)

for(i in 1:nsim){
  admit_star <- rbinom(n = length(m_reduced$y),  
                       size = 1, 
                       p = m_reduced$fitted.values)
  
  sim_data <- grad_app |>
    mutate(admit_star = admit_star)
  
  m_full_sim <- glm(admit_star ~ gpa + gre + as.factor(rank),
                    family = binomial, data = sim_data)
  
  parametric_boot_stats[i] <- summary(m_full_sim)$coefficients[3, 3]
}

hist(parametric_boot_stats, xlim = c(-3.5, 3.5))
```

.large[
.question[
What does this distribution look like?
]
]

---

### Activity

```{r echo=F, fig.width=8, fig.height=5}
hist(parametric_boot_stats, xlim = c(-3.5, 3.5))
```

.large[
.question[
How unusual is the observed test statistic of 2.07, compared to this null distribution?
]
]

---

### p-value

```{r echo=F, fig.width=8, fig.height=5}
hist(parametric_boot_stats, xlim = c(-3.5, 3.5))
```

.large[
```{r}
mean(abs(parametric_boot_stats) > 2.07)
```
]

---

### p-value

```{r}
mean(abs(parametric_boot_stats) > 2.07)

summary(m_full)$coefficients
```





