---
title: "Outliers and multicollinearity"
author: "Ciaran Evans"
output: beamer_presentation
---

## Motivating example

```{r fig.width=3.5, fig.height=2.5}
#| message: FALSE
#| warning: FALSE
#| echo: FALSE
#| 
library(robustbase)
library(tidyverse)
foodstamp |>
  ggplot(aes(x = ifelse(participation == 0, "no", "yes"), 
             y = log(income + 1))) +
  geom_boxplot() +
  theme_bw() +
  labs(x = "Participation in food stamp program", 
       y = "log(Income + 1)")
```

**Question:** Why are we concerned about outliers when we fit a regression model?

## Motivating example

```{r fig.width=3, fig.height=2}
#| message: FALSE
#| warning: FALSE
#| echo: FALSE
#| 
foodstamp |>
  ggplot(aes(x = ifelse(participation == 0, "no", "yes"), 
             y = log(income + 1))) +
  geom_boxplot() +
  theme_bw() +
  labs(x = "Participation in food stamp program", 
       y = "log(Income + 1)")
```

```{r, echo=F}
m1 <- glm(participation ~ log(income + 1), family = binomial,
          data = foodstamp)

summary(m1)$coefficients |>
  round(digits = 3) |>
  knitr::kable()
```

**Question:** How would the estimated coefficients change if the potential outlier were removed?

## Cook's distance

```{r, echo=F}
m1 <- glm(participation ~ log(income + 1), family = binomial,
          data = foodstamp)

summary(m1)$coefficients |>
  round(digits = 3) |>
  knitr::kable()

m2 <- glm(participation ~ log(income + 1), family = binomial,
          data = foodstamp[-5,])

summary(m2)$coefficients |>
  round(digits = 3) |>
  knitr::kable()
```

**Cook's distance:**

## Cook's distance

```{r, echo=F, fig.width = 4, fig.height = 3.5}
library(patchwork)
p1 <- foodstamp |>
  ggplot(aes(y = ifelse(participation == 0, "no", "yes"), 
             x = log(income + 1))) +
  geom_boxplot() +
  theme_bw() +
  labs(y = "Participation in \n food stamp program", 
       x = "log(Income + 1)")

p2 <- foodstamp |>
  mutate(inc = log(income + 1),
         cook = cooks.distance(m1)) |>
  ggplot(aes(x = inc, y = cook)) +
  geom_point(size = 0.75) +
  labs(x = "log(income + 1)", y = "Cook's distance") +
  theme_bw()

p1/p2
```

## Handling outliers

Should I just remove any unusual observations in the data?

\vspace{4cm}

## College scorecard data

The College Scorecard, compiled by the US Department of Education, is meant to be a tool for college applicants to compare different institutions. Variables include:

* `CONTROL`: whether the school is public or private
* `SATVRMID`: midpoint of SAT critical reading scores of students attending the school
* `SATMTMID`: midpoint of SAT math scores
* `SATWRMID`: midpoint of SAT writing scores
* `ACTCMMID`: midpoint of the ACT cumulative scores
* `ACTENMID`: midpoint of ACT English scores
* `ACTMTMID`: midpoint of ACT math scores
* `ACTWRMID`: midpoint of ACT writing scores
* `UGDS`: number of undergraduate students at the school
* `NPT4`: average cost to attend the school
* `RPY_3YR_70`: whether the proportion of students actively repaying their loans 3 years after graduation is > 70% (1 = yes, 0 = no)

## College scorecard data

```{r, echo=F, fig.width=3.5, fig.height=3.5, message=F, warning=F}
scorecard <- read_csv("https://sta112-s26.github.io/projects/scorecard.csv")
library(GGally)
scorecard %>%
  ggpairs(columns = 4:10,
          upper = list(continuous = wrap("points", size = 0.5)),
          lower = list(continuous = GGally::wrap(ggally_cor, stars = F))) +
  theme_classic() +
  theme(text = element_text(size = 5))
```

## Multicollinearity

**Multicollinearity:** (linear) relationship between 2 or more explanatory variables.

\vspace{5cm}

## Example: multicollinearity

With all the test score variables:

```{r}
#| echo: FALSE

m1 <- glm(RPY_3YR_70 ~ SATVRMID + SATMTMID + SATWRMID +
            ACTCMMID + ACTENMID + ACTMTMID + ACTWRMID,
          family = binomial, data = scorecard)

summary(m1)$coefficients |>
  round(digits = 3) |>
  knitr::kable()
```

With just one test score variable:

```{r}
#| echo: FALSE

m1 <- glm(RPY_3YR_70 ~ SATVRMID,
          family = binomial, data = scorecard)

summary(m1)$coefficients |>
  round(digits = 3) |>
  knitr::kable()
```

## Variance inflation factors

## Handling multicollinearity

```{r, include=F}
scorecard <- scorecard |>
  select(-c(INSTNM, STABBR, RPY_3YR_RT_SUPP, MD_EARN_WNE_P10,
            GRAD_DEBT_MDN_SUPP))
```


```{r}
m1 <- glm(RPY_3YR_70 ~ .,
          family = binomial, data = scorecard)

car::vif(m1)
```

What should we do when there is high multicollinearity?

\vspace{2cm}



## Summary

* Outliers:
    * Change our coefficient estimates
    * Identify with EDA before model fitting
    * Also check with Cook's distance after model fitting
    * Clear errors, or points from a different population of interest, can be removed. Otherwise, report results with and without outliers
    
\vspace{0.5cm}

* Multicollinearity:
    * Makes it hard to interpret meaning of a coefficient (how do we hold other variables constant??)
    * Increases variability of estimated coefficients
    * Identify with pairs plots, correlation matrices before model fitting
    * Also check with VIFs after model fitting
    * Options: use a subset of the explanatory variables; combine several explanatory variables (average, PCA, etc.); ignore (if only predicting)
    
## Class activity

* Work on the class activity in R (linked on Canvas and the course website)
* You are welcome to work in groups
* At the end of class, submit HTML file on Canvas (one per group, listing all group member names)


