---
title: "Parametric bootstrap and visual inference"
format: html
editor: source
---

**Instructions:** 

* Work with a neighbor to answer the following questions 
* To get started, download the [class activity template](https://sta214-f26.github.io/class_activities/ca_08_template.qmd) file
* When you are finished, render the file as an HTML and submit the HTML to Canvas (let me know if you encounter any problems)

## Setup

For this activity, you will need the [`glmhelpR`](https://github.com/ciaran-evans/glmhelpR) and the `statmod` packages installed. 

You should already have installed `glmhelpR` on the previous class activity. **If not**, then run the following code in your R console. If R says that the `pak` package is unavailable, uncomment the first line and try again.

```{r, eval=F}
# install.packages("pak")
pak::pak("ciaran-evans/glmhelpR")
```

To install the `statmod` package, use

```{r, eval=F}
install.packages("statmod")
```

in the console.


## Fitted model

Using the dengue data, we fit a model with WBC and create a randomized quantile residual plot:

```{r, message=F, warning=F}
# install.packages("statmod")

library(tidyverse)
library(glmhelpR)

m_dengue <- glm(Dengue ~ WBC, data = dengue, family = binomial)

data.frame(x = model.frame(m_dengue)$WBC, 
           residuals = statmod::qresid(m_dengue)) |>
  ggplot(aes(x = x, y = residuals)) +
  geom_point(size=0.5, alpha=0.5) +
  geom_smooth() +
  theme_bw() +
  labs(y = "Randomized quantile residuals", 
       x = "WBC")
```

**Our question:** Is the plot consistent with a correctly-specified model?

## Parametric bootstrap

**Step 1:** Generate new responses $Y_i^* \sim Bernoulli(\widehat{\pi}_i)$

1. **Question:** Run the `R` code to create the new simulated responses:

```{r, eval=F}
y_star <- rbinom(n = length(m_dengue$y), 
                 size = 1, 
                 p = m_dengue$fitted.values)
```


**Step 2:** *Re-fit* the logistic regression model with the data $(\mathbf{x}_1, Y_1^*),...,(\mathbf{x}_n, Y_n^*)$

2. **Question:** Fill in the `R` code to re-fit the model with the simulated responses:

```{r, eval=F}
sim_data <- data.frame(WBC = model.frame(m_dengue)$WBC,
                       Dengue = ...)

m_sim <- glm(..., data = sim_data, family = binomial)
```

**Step 3:** Make a new quantile residual plot from the fitted model on the *simulated* data.

3. **Question:** Fill in the `R` code to create a quantile residual plot from the simulated model:


```{r, eval=F}
data.frame(x = model.frame(m_sim)$WBC, 
           residuals = ...) |>
  ggplot(aes(x = x, y = residuals)) +
  geom_point(size=0.5, alpha=0.5) +
  geom_smooth() +
  theme_bw() +
  labs(y = "Randomized quantile residuals", 
       x = "WBC")
```

4. **Question:** How does new quantile residual plot compare to the original?

## Visual inference

We want to repeat the parametric bootstrap process multiple times, and compare the quantile residual plots for the original data and all the simulated datasets.

This can be done efficiently with the `regressinator` package. The following code will produce a *line-up* of 20 different quantile residual plots, in a *random order*. One of the plots is for the original data, while the other 19 are from parametric bootstrap samples.

```{r, eval=F}
# install.packages("regressinator")
library(regressinator)

qresid_wbc <- function(f){
  data.frame(resid = statmod::qresid(f),
             x = model.frame(f)$WBC)
}

model_lineup(m_dengue, fn = qresid_wbc, nsim = 20) |> 
  ggplot(aes(x = x, y = resid)) +
  geom_point(size = 0.5) +
  geom_smooth() +
  facet_wrap(vars(.sample)) +
  labs(x = "WBC", y = "Residual") +
  theme_bw()
```

5. **Question:** Can you identify which plot corresponds to the *original* data? If so, this suggests that our model might not be correctly specified!



