---
title: "Class activity: logistic regression diagnostics"
format: html
---

**Instructions:** 

* Work individually or with a neighbor to answer the following questions 
* To get started, download the [class activity template](https://sta214-f26.github.io/class_activities/ca_09_template.qmd) file
* When you are finished, render the file as an HTML and submit the HTML to Canvas (let me know if you encounter any problems)

## Setup

For this activity, you will need the `mvtnorm` and `GGally` packages installed. To install, run the following in the **console**:

```{r, eval=F}
install.packages("mvtnorm")
install.packages("GGally")
```


# Multicollinearity

In the first part of this class activity, we will explore the effects of multicollinearity on a fitted logistic regression model.

The following code generates three correlated explanatory variables. The correlation between the three variables, $\rho$, is specified by `rho` in the code. Simulating the explanatory variables uses the `rmvnorm(...)` function of the `mvtnorm` package; you may need to install the `mvtnorm` package.

We then simulate data $Y_i \sim Bernoulli(\pi_i)$, with 

$$\pi_i = \dfrac{e^{\beta_0 + \beta_1 X_{i,1} + \beta_2 X_{i,2} + \beta_3 X_{i,3}}}{1 + e^{\beta_0 + \beta_1 X_{i,1} + \beta_2 X_{i,2} + \beta_3 X_{i,3}}}$$

```{r, eval=F}
library(mvtnorm) # you will need to install mvtnorm first!

n <- 1000
rho <- 0
beta <- c(0, 1, 1, 1)

# Create the covariance matrix sigma
sigma <- diag(rep(1 - rho, 3)) + matrix(rep(rho, 9), nrow=3)

# Generate Xs from a multivariate normal distribution
X <- rmvnorm(n, sigma = sigma)

# make a design matrix X_complete which includes the initial column of 1s for 
# the intercept
X_complete <- cbind(1, X)

# generate probabilities p
p <- exp(X_complete %*% beta)/(1 + exp(X_complete %*% beta))

# generate bernoulli response y
y <- rbinom(n, 1, p)

# fit a model to all the explanatory variables
m1 <- glm(y ~ X, family = binomial)
summary(m1)
```

## Questions

1. Run the code above with $\rho = 0$ (`rho <- 0`). Then use the code below to check for multicollinearity in the data. Does multicollinearity appear to be a concern?

```{r, eval=F}
library(tidyverse)
library(GGally)

data.frame(X) |>
  ggpairs() +
  theme_classic()
```

2. When $\rho = 0$, how do the coefficient estimates compare to the *true* values in the simulation ($\beta_0 = 0, \ \beta_1 =  1, \ \beta_2 = 1, \ \beta_3 = 1)$)? Run the code several times to confirm.

3. Report the approximate standard errors of the estimated coefficients (from the `summary` output) when $\rho = 0$.

4. Now change the correlation $\rho$ to `rho = 0.6` in the simulation, and remake the pairs plot to check for multicollinearity. Does multicollinearity appear to be a concern?

5. Report the approximate standard errors of the estimated coefficients (from the `summary` output) when $\rho = 0.6$. How do these values compare to the results in question 3?

6. Finally, change the correlation $\rho$ to `rho = 1` in the simulation, and run the code. What changes?

# Influential points

In the second part of this class activity, we will explore the effects of influential points on our fitted regression model. The following code simulates data from the logistic regression model

$$Y_i \sim Bernoulli(\pi_i)$$

$$\log \left( \dfrac{\pi_i}{1 - \pi_i} \right) = -1 + 2 X_i$$
then adds a potential outlier at $X = -2$.

```{r, eval=F}
# simulate a single explanatory variable from a Normal distribution
x <- rnorm(100)

# create P(Y = 1 | X) for each entry in x
# Here log odds = -1 + 2x
p <- exp(-1 + 2*x)/(1 + exp(-1 + 2*x))

# Finally, simulate a binary response at each x
y <- rbinom(100, 1, p)

x1 <- c(x, -5)
y1 <- c(y, 1)

# Fit a model WITH the unusual observation
m1 <- glm(y1 ~ x1, family = binomial)
summary(m1)

# Fit a model WITHOUT the unusual observation
m2 <- glm(y ~ x, family = binomial)
summary(m2)
```

## Questions

7. Run the code several times. How does the potential outlier impact the parameter estimates?

8. One way to check for influential points is with *Cook's distance*. The code below creates a plot of Cook's distance for the model with the potential outlier. How does Cook's distance for the potential outlier ($X = -2$) compare to the other values of Cook's distance in the plot?

```{r, eval=F}
plot(x1, cooks.distance(m1))
```


9. Play with the location of the potential outlier (e.g., try $X = -5$, or $X = -10$). Then re-run the code several times. How does the location of the outlier change its influence on the parameter estimates? What about the value of Cook's distance?

