---
title: "Class activity: bootstrap resampling"
format: html
---

**Instructions:** 

* Work individually or with a neighbor to answer the following questions 
* To get started, download the [class activity template](https://sta214-f26.github.io/class_activities/ca_13_template.qmd) file
* When you are finished, render the file as an HTML and submit the HTML to Canvas (let me know if you encounter any problems)

## Dengue data

Examining the dengue data, we create boxplots showing the difference in WBC between positive and negative dengue cases:

```{r message=F, warning=F}
library(tidyverse)

dengue <- read_csv("https://sta214-f26.github.io/homework/dengue.csv")

dengue |>
  ggplot(aes(x = as.factor(Dengue), y = WBC)) +
  geom_boxplot() +
  theme_bw() +
  labs(x = "Dengue")
```

To measure the difference in WBC between the two groups, we can compare the means from each group:

$$\overline{WBC}_1 - \overline{WBC}_0$$

Computing this difference in R:

```{r}
dengue |>
  summarize(mean_diff = mean(WBC[Dengue == 1]) - mean(WBC[Dengue == 0])) |>
  as.numeric() # convert the result to a number, not data frame
```

Of course, if we collected another sample and computed the difference in means again, we would get a different value!

**Our question:** how much does $\overline{WBC}_1 - \overline{WBC}_0$ **vary from sample to sample?**

## Bootstrap resampling

We typically can't collect a new sample of data. However, we can **resample** (with replacement) from our *existing* sample of data. This process is called **bootstrap resampling**.

### Step 1: Resample with replacement

To create our bootstrap sample, we will sample with replacement from the rows of the dengue data.

1. Run the following code to generate a bootstrap sample.

```{r}
dengue_boot <- dengue |>
  sample_n(size = n(), replace=T)
```

### Step 2: Re-compute the statistic of interest

Now we need to recompute the statistic of interest, in this case the difference in means.

2. Re-make the boxplots with your bootstrap sample (`dengue_boot`). Does the boxplot look similar or different to the plot created from the original data?

3. Re-calculate the difference in means ($\overline{WBC}_1 - \overline{WBC}_0$), using data from the bootstrap sample (`dengue_boot`).

### Step 3: Repeating many times!

Just one bootstrap sample hardly helps us estimate variability. We need many bootstrap samples!

To do this, we will use a programming tool called a **loop**. In a loop, we repeat the same process a specified number of times.

The code below provides the general structure for a loop in R. In this case, we will repeat the whole bootstrap process `n_boot = 1000` times.

```{r, eval=F}
# desired number of bootstrap samples
n_boot = 1000

# create a vector in which to store the bootstrap samples
boot_ests <- rep(NA, n_boot)

# do the bootstrapping
for(i in 1:n_boot){
  # fill in the code to create a bootstrap sample 
  # (question 1)
  ...
  
  # fill in the code to re-calculate the difference in means
  # (question 3). We will store the difference in the
  # boot_ests vector
  boot_ests[i] <- ...
}

# make a histogram of the results
hist(boot_ests)
```

```{r, include=F}
# desired number of bootstrap samples
n_boot = 1000

# create a vector in which to store the bootstrap samples
boot_ests <- rep(NA, n_boot)

# do the bootstrapping
for(i in 1:n_boot){
  # fill in the code to create a bootstrap sample 
  # (question 1)
  dengue_boot <- dengue |>
    sample_n(size = n(), replace=T)
  
  # fill in the code to re-calculate the difference in means
  # (question 3). We will store the difference in the
  # boot_ests vector
  boot_ests[i] <- dengue_boot |>
    summarize(mean_diff = mean(WBC[Dengue == 1]) - 
                          mean(WBC[Dengue == 0])) |>
    as.numeric()
}

# make a histogram of the results
hist(boot_ests)
```

4. Fill in the code (replace the `...` with appropriate code from questions 1 and 3), and run it.

5. Compute the standard deviation of the `boot_ests`. This is our estimate of the standard error of the difference in means, $SE(\overline{WBC}_1 - \overline{WBC}_0)$.

6. Look at the shape of the histogram. What does that distribution look like?

### Step 4: using the results

Now let's use the bootstrap results to get a 95% confidence interval for the true difference in means ($\mu_1 - \mu_0$).

There are multiple ways we could try to do this. We will look at the **bootstrap percentile interval** and the **Normal interval with bootstrap standard errors**.

#### Bootstrap percentile interval

A 95% bootstrap percentile interval is the range from the 2.5th percentile to the 97.5th percentile (the middle 95% of the bootstrap samples).

7. Run the code below in R to get the percentile interval:

```{r, eval=F}
quantile(boot_ests, c(0.025, 0.975))
```


#### Normal interval with bootstrap SE

The Normal interval is based on the fact that the bootstrap distribution has an approximately normal distribution; we just use the bootstrap to estimate the standard error. In this case, the bootstrap interval is

$$(\overline{WBC}_1 - \overline{WBC}_0) \pm z^* SE(\overline{WBC}_1 - \overline{WBC}_0)$$

* We calculated the first piece, $\overline{WBC}_1 - \overline{WBC}_0$, at the start: -4.36.

* We estimated the standard error in question 5.

* $z^*$ is the critical value for the desired confidence level (in this case, we want a 95% confidence level). For a 95% CI, $z^* = 1.96$

8. Using this information, calculate the Normal bootstrap interval and compare with the percentile interval.




