---
title: "Class activity: parametric bootstrap null distribution"
format: html
callout-icon: false
---

**Instructions:** 

* Work individually or with a neighbor to answer the following questions 
* To get started, download the [class activity template](https://sta214-f26.github.io/class_activities/ca_15_template.qmd) file
* When you are finished, render the file as an HTML and submit the HTML to Canvas (let me know if you encounter any problems)

## Setup

In this activity, you will work with data on 400 applicants to graduate school. For each applicant, we have:

* `admit`: whether they were admitted to a graduate program (0 = no, 1 = yes)
* `gre`: their GRE score (out of 800)
* `gpa`: their undergraduate GPA (out of 4.0)
* `rank`: a proxy for school prestige (1 = most prestigious, 4 = least prestigious)

The data can be downloaded in R with:

```{r}
grad_app <- read.csv("https://stats.idre.ucla.edu/stat/data/binary.csv")
```

## Previously

Previously, we used the parametric bootstrap to approximate the null distribution for the $Z$ statistic for testing a single coefficient in a logistic regression model.

Here is the full code from the previous activity:

```{r, eval = F}
library(tidyverse)

m_full <- glm(admit ~ gpa + gre + as.factor(rank),
              family = binomial, data = grad_app)

m_reduced <- glm(admit ~ gpa + as.factor(rank),
                 family = binomial, data = grad_app)

nsim <- 5000
parametric_boot_stats <- rep(NA, nsim)

for(i in 1:nsim){
  admit_star <- rbinom(n = length(m_reduced$y),  
                       size = 1, 
                       p = m_reduced$fitted.values)
  
  sim_data <- grad_app |>
    mutate(admit_star = admit_star)
  
  m_full_sim <- glm(admit_star ~ gpa + gre + as.factor(rank),
                    family = binomial, data = sim_data)
  
  parametric_boot_stats[i] <- summary(m_full_sim)$coefficients[3, 3]
}

# approximate a p-value from parametric bootstrap
# 2.07 was the observed difference in the original data
mean(abs(parametric_boot_stats) > 2.07)
```

Now we want to use parametric bootstrapping again, but this time to compare the deviance of full and reduced models (allowing us to potentially test hypotheses about *multiple* parameters simultaneously!)

## Research question

We decide to use the logistic regression model

$$
\begin{aligned}
\text{Admit}_i &\sim Bernoulli(\pi_i) \\
\log \left( \dfrac{\pi_i}{1 - \pi_i} \right) &= \beta_0 + \beta_1 \text{GPA}_i + \beta_2 \text{GRE}_i + \\
& \hspace{0.5cm} \beta_3 \text{Rank2}_i + \beta_4 \text{Rank3}_i + \beta_5 \text{Rank4}_i
\end{aligned}
$$

**Research question:** Is there a relationship between school rank and grad school acceptance, after accounting for GPA and GRE score?

### Hypothesis testing

To answer this research question, we want to test the hypotheses

$H_0: \beta_3 = \beta_4 = \beta_4 = 0$

$H_A: \text{ at least one of } \beta_3, \beta_4, \beta_5 \neq 0$

These hypotheses give us the full model and the reduced model:

**Full model**:

$$
\begin{aligned}
\log \left( \dfrac{\pi_i}{1 - \pi_i} \right) &= \beta_0 + \beta_1 \text{GPA}_i + \beta_2 \text{GRE}_i + \\
& \hspace{0.5cm} \beta_3 \text{Rank2}_i + \beta_4 \text{Rank3}_i + \beta_5 \text{Rank4}_i
\end{aligned}
$$

**Reduced model**:

$$
\begin{aligned}
\log \left( \dfrac{\pi_i}{1 - \pi_i} \right) &= \beta_0 + \beta_1 \text{GPA}_i + \beta_2 \text{GRE}_i
\end{aligned}
$$

### Test statistic

Our test statistic is

$$G = \text{deviance}_{\text{reduced}} - \text{deviance}_{\text{full}}$$

Calculating this difference in R:

```{r}
m_full <- glm(admit ~ gpa + gre + as.factor(rank),
              family = binomial, data = grad_app)

m_reduced <- glm(admit ~ gpa + gre,
                 family = binomial, data = grad_app)

m_reduced$deviance - m_full$deviance
```

We then want to know: **how unusual is this test statistic, if in fact the null hypothesis is actually true?**

## Parametric bootstrapping

To determine whether the observed value of $G$ is unusual under $H_0$, we need the **null distribution** of our test statistic. We will use **parametric bootstrapping** to assess whether $G = 21.83$ is consistent with $H_0$, or is unusual.

### Step 1: Fit a model that assumes $H_0$ is true

**If** $H_0$ is true, then we don't need school rank in the model. We have already fit this reduced model (the `m_reduced` in the section immediately above).

### Step 2: Use the fitted model to generate new responses

Now we will use our reduced model to generate new responses $\text{Admit}_1^*,...,\text{Admit}_n^*$. Because these new responses are generated from our reduced model, we know that they have no relation with school rank!

::: {.callout-tip}
## Question 1

Use the code below to generate the new responses:

```{r, eval=F}
admit_star <- rbinom(n = length(m_reduced$y), 
                     size = 1,
                     p = m_reduced$fitted.values)

sim_data <- grad_app |>
  mutate(admit_star = admit_star)
```

:::

### Step 3: *Re-fit* the full model with the simulated responses

Our simulated data (`sim_data`) now contains a response (`admit_star`) that is created to be unrelated to GRE score -- that is, we have forced $H_0$ to be true! To see how the difference in deviance behaves in this situation, we need to **re-fit** both the **full** and **reduced** models on this simulated data.

::: {.callout-tip}
## Question 2

Re-fit the full model on the simulated data (`sim_data`), and call the new fitted model `m_full_sim`. You should use GPA, GRE, and school rank as the explanatory variables, and $\text{Admit}^*$ (`admit_star`) as the response.

```{r, eval=F}
m_full_sim <- glm(...,
                  family = binomial, data = sim_data)
```

:::

::: {.callout-tip}
## Question 3

Re-fit the reduced model on the simulated data (`sim_data`), and call the new fitted model `m_reduced_sim`. You should use GPA and GRE as the explanatory variables, and $\text{Admit}^*$ (`admit_star`) as the response.

```{r, eval=F}
m_reduced_sim <- glm(...,
                  family = binomial, data = sim_data)
```

:::

### Step 4: Recompute the test statistic

::: {.callout-tip}
## Question 4

Recompute the change in deviance from the models you fit in questions 2 and 3 (`m_full_sim` and `m_reduced_sim`). How does the test statistic compare to value computed on the original data (21.83)?
:::

### Step 5: Repeat many times!

We want a *distribution* for $G$ under $H_0$, so we need to repeat many times. 

::: {.callout-tip}
## Question 5

Fill in the code below (all the `...`) to repeat the code from questions 1--4 many times:

```{r, eval=F}
nsim <- 5000
parametric_boot_stats <- rep(NA, nsim)

for(i in 1:nsim){
  # create a new parametric bootstrap sample
  # (code from question 1)
  ...
  
  # re-fit the full and reduced models on the 
  # simulated data
  # (code from questions 2 - 3)
  ...
  
  # recompute the change in deviance, and save
  parametric_boot_stats[i] <- ...
}

hist(parametric_boot_stats)
```


What does the null distribution of $G$ look like? Is a value of 21.83 unusual for this distribution?

:::

### Step 6: Approximate a p-value

Finally, we will use the parametric bootstrap null distribution to approximate a p-value for our original test statistic. 

Remember that a p-value measures the probability of a value "as or more extreme" as the observed test statistic. In our case, a bigger change in deviance is more extreme.

In other words, we want $P(G > 21.83 \ | \ H_0)$.

::: {.callout-tip}
## Question 6

The following code will calculate the p-value:

```{r, eval=F}
mean(parametric_boot_stats > 21.83)
```

Report the p-value. Would you reject or fail to reject our null hypothesis?

:::



