---
title: "Class activity: parametric bootstrap null distribution"
format: html
callout-icon: false
---

**Instructions:** 

* Work individually or with a neighbor to answer the following questions 
* To get started, download the [class activity template](https://sta214-f26.github.io/class_activities/ca_14_template.qmd) file
* When you are finished, render the file as an HTML and submit the HTML to Canvas (let me know if you encounter any problems)

## Setup

In this activity, you will work with data on 400 applicants to graduate school. For each applicant, we have:

* `admit`: whether they were admitted to a graduate program (0 = no, 1 = yes)
* `gre`: their GRE score (out of 800)
* `gpa`: their undergraduate GPA (out of 4.0)
* `rank`: a proxy for school prestige (1 = most prestigious, 4 = least prestigious)

The data can be downloaded in R with:

```{r}
grad_app <- read.csv("https://stats.idre.ucla.edu/stat/data/binary.csv")
```

## Research question

We decide to use the logistic regression model

$$
\begin{aligned}
\text{Admit}_i &\sim Bernoulli(\pi_i) \\
\log \left( \dfrac{\pi_i}{1 - \pi_i} \right) &= \beta_0 + \beta_1 \text{GPA}_i + \beta_2 \text{GRE}_i + \\
& \hspace{0.5cm} \beta_3 \text{Rank2}_i + \beta_4 \text{Rank3}_i + \beta_5 \text{Rank4}_i
\end{aligned}
$$

We fit the model in R:

```{r}
m_full <- glm(admit ~ gpa + gre + as.factor(rank),
              family = binomial, data = grad_app)

summary(m_full)
```


**Research question:** Is there a relationship between GRE and grad school acceptance, after accounting for GPA and school rank?

### Hypothesis testing

To answer this research question, we want to test the hypotheses

$$H_0: \beta_2 = 0 \hspace{1cm} H_A: \beta_2 \neq 0$$

Our test statistic is

$$Z = \frac{\widehat{\beta}_2 - 0}{SE_{\widehat{\beta}_2}}$$

Using the output of the fitted model above,

$$Z = \frac{0.002264}{0.001094} = 2.07$$

We then want to know: **how unusual is this test statistic, if in fact the null hypothesis is actually true?**

## Parametric bootstrapping

To determine whether the observed value of $Z$ is unusual under $H_0$, we need the **null distribution** of our test statistic. We will use **parametric bootstrapping** to assess whether $Z = 2.07$ is consistent with $H_0$, or is unusual.

### Step 1: Fit a model that assumes $H_0$ is true

**If** $H_0$ is true, then we don't need $\text{GRE}$ in the model

::: {.callout-tip}
## Question 1

Fit another logistic regression model to the `grad_app` data, and call this new model `m_reduced`. (This is a *reduced* model because it is smaller than the *full* model that we fit above). Your reduced model should include GPA and school rank, but *not* GRE score.

```{r eval=F}
m_reduced <- glm(...,
                 family = binomial, data = grad_app)
```


:::

### Step 2: Use the fitted model to generate new responses

Now we will use our reduced model to generate new responses $\text{Admit}_1^*,...,\text{Admit}_n^*$. Because these new responses are generated from our reduced model, we know that they have no relation with GRE score!

::: {.callout-tip}
## Question 2

Use the code below to generate the new responses:

```{r, eval=F}
admit_star <- rbinom(n = length(m_reduced$y), 
                     size = 1,
                     p = m_reduced$fitted.values)

sim_data <- grad_app |>
  mutate(admit_star = admit_star)
```

:::

### Step 3: *Re-fit* the full model with the simulated responses

Our simulated data (`sim_data`) now contains a response (`admit_star`) that is created to be unrelated to GRE score -- that is, we have forced $H_0$ to be true! To see how $\widehat{\beta}_2$ and the test statistic $Z$ behave in this situation, we need to **re-fit** the **full** model on this simulated data.

::: {.callout-tip}
## Question 3

Re-fit the full model on the simulated data (`sim_data`), and call the new fitted model `m_full_sim`. You should use GPA, GRE, and school rank as the explanatory variables, and $\text{Admit}^*$ (`admit_star`) as the response.

```{r, eval=F}
m_full_sim <- glm(...,
                  family = binomial, data = sim_data)
```

:::

### Step 4: Recompute the test statistic

::: {.callout-tip}
## Question 4

Recompute $Z = \dfrac{\widehat{\beta}_2}{SE_{\widehat{\beta}_2}}$ from the model you fit in question 3 (`m_full_sim`). How does the test statistic compare to value computed on the original data (2.07)?
:::

### Step 5: Repeat many times!

We want a *distribution* for $Z$ under $H_0$, so we need to repeat many times. The following code will approximate the null distribution of $Z$ with 5000 parametric bootstrap samples, and make a histogram:

```{r, eval=F}
nsim <- 5000
parametric_boot_stats <- rep(NA, nsim)

for(i in 1:nsim){
  admit_star <- rbinom(n = length(m_reduced$y),  
                       size = 1, 
                       p = m_reduced$fitted.values)
  
  sim_data <- grad_app |>
    mutate(admit_star = admit_star)
  
  m_full_sim <- glm(admit_star ~ gpa + gre + as.factor(rank),
                    family = binomial, data = sim_data)
  
  parametric_boot_stats[i] <- summary(m_full_sim)$coefficients[3, 3]
}

hist(parametric_boot_stats)
```

::: {.callout-tip}
## Question 5

What does the null distribution of $Z$ look like? Is a value of 2.07 unusual for this distribution?

:::

### Step 6: Approximate a p-value

Finally, we will use the parametric bootstrap null distribution to approximate a p-value for our original test statistic. 

Remember that a p-value measures the probability of a value "as or more extreme" as the observed test statistic. In our case, the alternative hypothesis is $H_A: \beta_2 \neq 0$, so a value "as or more extreme" could be either $> 2.07$ or $< -2.07$.

In other words, we want $P(|Z| > 2.07 \ | \ H_0)$.

::: {.callout-tip}
## Question 6

The following code will calculate the p-value:

```{r, eval=F}
mean(abs(parametric_boot_stats) > 2.07)
```

Report the p-value. Would you reject or fail to reject our null hypothesis $H_0: \beta_2 = 0$?

:::



