---
title: "Information packet"
format: 
  pdf:
    header-includes:
      - \usepackage{float}
      - \floatplacement{table}{H}
editor: source
---

\vspace{-2.5cm}

```{r setup, include=F}
library(tidyverse)
library(patchwork)
library(glmhelpR)

ct <- read_csv("https://cmustatistics.github.io/data-repository/data/core-temperature.csv")
```

# Variable descriptions

| Variable | Description |
|----|-------------|
| YEAR | Year of surgery |
| Age | Patient's age at the time of the surgery |
| FEMALE | Patient's sex at birth (1 = female, 0 = male) |
| BMI | Patient's body mass index (BMI) at time of surgery |
| CharlsonScore | Patient's Charlson Comorbidity Index score (higher numbers indicate more comorbidities that could lead to death) |
| SurgeryType | This was meant to record the type of surgery (such as colostomy, ileostomy, or other procedures)|
| CHF | Does the patient have a history of congestive heart failure? (1 = yes, 0 = no) |
| VALVE | Does the patient have a history of peripheral vascular disease? (1 = yes, 0 = no) |
| RENLFAIL | Does the patient have a history of renal failure? (1 = yes, 0 = no) |
| LIVER | Does the patient have a history of liver disease? (1 = yes, 0 = no) |
| METS | Does the patient have cancer that has metastasized? (1 = yes, 0 = no) |
| TUMOR | Does the patient have a solid tumor that has *not* metastasized? (1 = yes, 0 = no) |
| COAG | Does the patient have coagulopathy (a blood clotting disorder)? (1 = yes, 0 = no) |
| OBESE | Is the patient obese? (Definition not specified; 1 = yes, 0 = no) |
| LYTES | Does the patient have fluid or electrolyte disorders? (1 = yes, 0 = no) |
| BLDLOSS | Does the patient have chronic blood loss anemia? (1 = yes, 0 = no) |
| ANEMDEF | Does the patient have a deficiency anemia? (1 = yes, 0 = no) |
| DRUG | Does the patient abuse drugs? (1 = yes, 0 = no) |
| SteroidHx | Did the patient use steroid drugs before surgery? (1 = yes, 0 = no) |
| ImmunosuppressantHx | Did the patient use immunosuppressive drugs before surgery? (1 = yes, 0 = no) |
| SurgDuration | Duration of the surgery, in minutes |
| Open | Type of surgery. 1 = open surgery, 0 = laparoscopic ("minimally invasive") surgery. |
| DelayedHealing | Was the patient's healing after surgery delayed? (1 = yes, 0 = no) |
| TWATemp | Time weighted average of the patient's core temperature during surgery, °C |
| LastReadingTemp | Last recorded core temperature prior to surgery, °C |
| EndCaseTemp | Patient's core temperature at the end of surgery, °C |
| SeriousInfection | Did the patient develop a serious infection within 30 days of surgery? This includes the infections in the SSIDeep(fascia), SSIOrganSpace, AbsessIntraAb, AbsessPelvic, Cdiff, Pneumonia, Pneumonia(aspiration), and Sepsis variables. (1 = yes, 0 = no) |
| SuperficialInfection | Did the patient develop a superficial infection within 30 days of surgery? This includes the infections in the SSISuperficial(skin), WoundInfection, and FascialDehiscence variables, as well as perineal wound problems. (1 = yes, 0 = no) |
| DurationHosp | Time the patient stayed in the hospital after surgery, in days |
| LOS | Time the patient stayed in the hospital, including time before surgery, in days |
| DEAD | Did the patient die in the hospital? (1 = yes, 0 = no) |


\newpage

# Empirical logit plots

```{r, echo=F, fig.width=9, fig.height=6, message=F}
p1 <- ct |> 
  emp_logit_plot(SeriousInfection ~ TWATemp,
                 num_bins = 15, pt_size = 0.5) +
  geom_smooth(color = "orange")

p2 <- ct |> 
  emp_logit_plot(SeriousInfection ~ SurgDuration,
                 num_bins = 15, pt_size = 0.5) +
  geom_smooth(color = "orange")

p3 <- ct |> 
  emp_logit_plot(SeriousInfection ~ CharlsonScore,
                 num_bins = 15, pt_size = 0.5) +
  geom_smooth(color = "orange")

p4 <- ct |> 
  emp_logit_plot(SeriousInfection ~ Age,
                 num_bins = 15, pt_size = 0.5) +
  geom_smooth(color = "orange")

p5 <- ct |> 
  emp_logit_plot(SeriousInfection ~ BMI,
                 num_bins = 15, pt_size = 0.5) +
  geom_smooth(color = "orange")

(p1 + p2 + p3)/(p4 + p5 + plot_spacer())
```

\newpage

# Summary of variables

```{r, echo=F}
ct |>
  select(YEAR, Age, FEMALE, SurgeryType, BMI,
         CharlsonScore, CHF, VALVE, RENLFAIL, LIVER, METS, TUMOR, COAG, OBESE, LYTES, SurgDuration, Open, TWATemp, LastReadingTemp, EndCaseTemp, SeriousInfection,
         SuperficialInfection, DurationHosp, LOS, DEAD) |>
  gtsummary::tbl_summary()
```















