---
title: "Class activity: logistic regression predictions"
format: html
---

**Instructions:** 

* Work individually or with a neighbor to answer the following questions 
* To get started, download the [class activity template](https://sta214-f26.github.io/class_activities/ca_11_template.qmd) file
* When you are finished, render the file as an HTML and submit the HTML to Canvas (let me know if you encounter any problems)

## SBA data

The US Small Business Administration (SBA) helps guarantee loans from small businesses. We are interested in predicting whether a business is likely to *default* on their loan (fail to repay the loan), using some information about the loan and the business.

* `Default`: whether the business defaulted on the loan (1 = yes, 0 = no)
* `UrbanRural`: 1 if business is in urban area, 2 if business is in rural area, 0 if unknown
* `NewExist`: 1 if business already existed, 2 if business is new, 0 if unknown
* `Term`: Length of the loan term (months)
* `NoEmp`: Number of employees of the business before receiving the loan
* `GrAppv`: Gross amount of loan approved by the bank, in dollars
* `SBA_Appv`: Amount of the loan guaranteed by the SBA, in dollars
* `DisbursementGross`: The amount of money disbursed (loaned), in dollars

You can load the SBA data into R with

```{r, eval=F}
sba <- read.csv("https://sta214-f26.github.io/class_activities/sba_small.csv")
```

### Questions

1. Fit a logistic regression model to predict loan default using loan term, log(disbursement gross), and whether the business is urban or rural:

```{r, eval=F}
m1 <- glm(...)
```


2. Use the following code to create a confusion matrix for the SBA predictions (replace `m1` with the name of your fitted model, if you called it something different):

```{r, eval=F}
table(Prediction = m1$fitted.values > 0.5, 
      Truth = sba$Default)
```

3. Calculate accuracy, sensitivity, specificity, and positive predictive value.

4. Is our model doing a good job at predicting loan default?

5. What happens we change the threshold? Instead of a threshold of 0.5, experiment with other values (e.g., 0.3, 0.7, etc.)



